Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-21 thesis/paper_draft.qmd
Refine tourism section (and slight changes to others) from George feedback

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g., the sum of all countries' sales forecasts does not equal the total sales forecast). Such forecasts are called *incoherent*. Incoherence undermines downstream decisions that require internal consistency and can degrade forecasting performance. To tackle this problem, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints). It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation. Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs. @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation. Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance. MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions. MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].

@@

## Exploring Effects of Hierarchy's Size

@@

Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across the two structures (small and large) and time series lengths (short and long). MinT with shrinkage (*MinT-S*) and its variants (as defined in @tbl-cov) are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants (as defined in same table) are in purple. The solid lines represent vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g. *MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g. *MinT-S(SV)*, *MinT-N(hcov)*).

@@

Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which splits into the seven states and territories. Each state is further sub-divided into tourism zones, which in turn break down into 77 regions. A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement. Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate. @tbl-tourism-structure depicts this structure.

@@

The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.

@@

@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series. Post-processing reconciliation approaches have the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.

@@

To assess forecasting performance between models, we adopt a rolling-window cross-validation scheme. Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches. To estimate the NOVELIST and its variants, we would have an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv. We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015. In total, this results in 97 out-of-sample windows. The entire procedure is illustrated in @fig-tourism-cv.

@@

```{r}
#| label: fig-tourism-MSE-heat
#| fig.height: 5
#| fig.width: 7
#| out.width: "90%"
#| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts. The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star."

@@

@fig-tourism-MSE-line show a main difference compared to the one of simulation results. Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts. Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.

@fig-tourism-MSE-heat provides a complementary heatmap view, scaling each method's MSE improvement to a 0-100 range for better visual discrimination. Here, *MinT-N(PC1)* and *MinT-S(PC1)* emerge as the top performers across most horizons. The heatmap underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments.

@@

Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* (purple bars) consistently outperforms *MinT-S* (green bars) across univariate and multivariate scores. The PC1-adjusted variants again yield improvements over their vanilla counterparts, with *MinT-N(PC1)* leading overall. In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods. One surprising finding is that all MinT variants underperform the base forecasts when moving to the 95% Winkler score, while *OLS* performs best.

@@

To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* underperforms the base at all levels, especially in higher aggregated levels. From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside. The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

```{r}
#| label: fig-tourism-radar
#| fig.height: 7
#| fig.width: 7
#| out.width: "70%"
#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler score at 80% and 95% intervals, CRPS, and Energy) over the base forecasts. The scores are scaled to a range of 0 to 100, with larger values indicating better performance. The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0). Only the top 4 MinT approaches are shown (with the base model)."

@@

The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, Winkler score at 80% and 95% intervals, and Energy, extending the improvements seen in the single-metric panels. Except for the Winkler score at 95% interval, all MinT variants outperform the base forecasts.

>>>>

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g., the sum of all countries' sales forecasts does not equal the total sales forecast). Such forecasts are called *incoherent*. Incoherence undermines downstream decisions that require internal consistency and can degrade forecasting performance. To tackle this problem, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints). It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation. Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs. @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.

Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance. MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions. @Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series. Post-processing reconciliation has the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT. MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].

@@

## Exploring Effects of Hierarchy's Size {#sec-sim-hierarchy-size}

@@

Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across the two structures (small and large) and time series lengths (short and long). We evaluate ten different MinT variants based on the covariance estimators discussed in @tbl-cov, with visual distinctions made through colors, line types, and point shapes. MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple. The solid lines represent vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g. *MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g. *MinT-S(SV)*, *MinT-N(hcov)*).

@@

<!-- briefly talk about the important of forecasting tourism when describing the data-->

Forecasting domestic tourism flows is of great economic and policy significance, as policymakers and stakeholders rely on accurate forecasts to make informed decisions regarding resource allocation, infrastructure development, and marketing strategies. The domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which splits into the seven states and territories. Each state is further sub-divided into tourism zones, which in turn break down into 77 regions. A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement. Intersecting this geographic hierarchy is a second dimension: policymakers are also interested in travel motives. This partitions tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped structure of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate. @tbl-tourism-structure depicts this structure.

@@

The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series. This gives a challenging high-dimensional forecasting problem with number of series (560) doubled the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks Corporation sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.

@@

@@

To assess forecasting performance between models, we adopt a rolling-window cross-validation scheme. Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches. To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv. We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015. In total, this results in 97 out-of-sample windows. The entire procedure is illustrated in @fig-tourism-cv.

@@

```{r}
#| label: fig-tourism-MSE-heat
#| fig.height: 5
#| fig.width: 7
#| out.width: "90%"
#| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts. The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star. This re-scaling does not affect relative rankings among methods."

@@

<!--
- describe the fig
- say models in @tbl-cov
- encoded by color, linetype, shape similar to simulation plot
- quickly refresh the memory of what each means
-->

@fig-tourism-MSE-line reports percentage relative improvements in MSE of reconciled forecasts over the incoherent base ARIMA forecasts, across horizons from 1 to 12 months. We evaluate ten MinT variants (as defined in @tbl-cov), similar to the simulation's @fig-sim-results-1. The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple. Solid lines represent "vanilla" *MinT-S* and *MinT-N*; dashed lines with dot points denote the PC-adjusted variants (e.g. *MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g. *MinT-S(SV)*, *MinT-N(hcov)*).

Three patterns stand out. First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small. Second, variants that alter the multi-step covariance, either via scaled variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis. Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods. Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons. However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.

<!-- @fig-tourism-MSE-line shows a main difference compared to the one of simulation results. Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts. Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.  -->

@fig-tourism-MSE-heat complements the line plot with a heatmap that standardises each method’s MSE improvement to a 0–100 scale to enhance visual discrimination across horizons. Darker shades indicate larger improvements. The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers). Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and h-step residuals).

@@

Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* variants (purple bars) consistently outperform *MinT-S* variants (green bars) across univariate and multivariate scores. The PC1-adjusted variants again yield improvements over their vanilla counterparts, with *MinT-N(PC1)* leading overall. In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods. One surprising finding is that all MinT variants produce inferior forecasts compared to the base forecasts when considering 95% Winkler score, while *OLS* performs best.

@@

To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* produces inferior forecasts compared to the base forecasts at all levels, especially in higher aggregated levels. From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals. The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and resulting in poor Winkler scores. *MinT-N* has relatively good coverage and improved Winkler score at bottom levels, but still is inferior at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

```{r}
#| label: fig-tourism-radar
#| fig.height: 7
#| fig.width: 7
#| out.width: "70%"
#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler score at 80% and 95% intervals, CRPS, and Energy) over the base forecasts. The scores are scaled to a range of 0 to 100, with larger values indicating better performance. The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0). Only the top 4 MinT approaches are shown, together with the base forecasts."

@@

The summary radar graph in @fig-tourism-radar consolidates these findings. Among the selected MinT variants considering probabilistic criteria, *MinT-N(PC1)* (yellow polygon) clearly leads across CRPS, Winkler score at 80% and 95% intervals, and Energy score, extending the improvements seen in the single-metric panels. Except for the Winkler score at 95% interval, all MinT variants outperform the base forecasts.
